---
title: "2021-09-03_COVID_prolonged_positive_manuscript_tables_figures"
author: "Shawn Hawken"
date: "9/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
#knitr::opts_chunk$set(dev = c('pdf'),fig.path='../../manuscripts/prolonged_positive_manuscript/figures_and_tables/')
rmarkdown::render
```
# Clinical characteristics of patients with prolonged positivity with SARS-CoV2  
-tables and figures for manuscript

### Load workspace image containing data to generate figures and table for manuscript
-data processed and saved according to date of file
-load in de-identified data  

```{r, read_in_data}
# pathology drive: 
#setwd('/Volumes/Clin Micro Miller Research/Shawn_COVID_research/COVID/COVID_scripts')

load(file="2021-09-03_prolonged_positive_de-identified_workspace.R")

# load packages
library(ggplot2)
library(reshape2)
library(knitr)
library(lubridate)
library(kableExtra)
library(openxlsx)
library(ggrepel)
library(ggforce)
library(ape)
library(msa)
library(gtsummary)
library(tidyr)
library(table1)
#library(boot)


#custom function libraries for analysis and generating figures
source("lib/2020-11-18_functions_for_repeat_testing_analysis.R")
```

## Results section for manuscript

__Summary of SARS-CoV2 testing during study__
Total tests run  
Total positive and total negatives  
Distribution of tests across platforms used 
Distribution of specimen types tested   
Total patients tested  
Total patients tested >1 time  
Total patients with >1 positive test  
Distribution of time between positive tests for patients with >1 positive test  
Total patients with first and last positive tests w/in 1 incubation period (14 days)  
Total patients with first and last positive test >=21 days apart  
total patients with positive/negative/positive pattern


```{r, summarize_dataset}

### Summarize number of unique total tests run at UNC hospital and Collected through McLendon labs:

# print min and max dates for current dataset
print("Date range of Epic data for study: ")
print(min(covid.df$Collected, na.rm = T))
print(max(covid.df$Collected,na.rm = T))

print('Total tests run:')
print(dim(covid.df)[1])

print("breakdown of negative and positive tests overall during timeframe: ")
print(table(covid.df$pos.test))


print("Distribution of tests across platforms used: ")
print(table(covid.df$platform))



print("Distribution of specimen types for tests performed: ")
print(sort(table(as.vector(covid.df$specimen.Type.clean))))
print(sort((table(as.vector(covid.df$specimen.Type.clean))/dim(covid.df)[1])*100))


print("Distribution of positive and negative tests by specimen types for tests performed: ")
#summary table positive and negative tests by speciment type
pos.neg.spec.type.df<-as.data.frame(do.call(cbind, by(as.vector(covid.df$specimen.Type.clean), covid.df$pos.test, summary)))
# add % positive column
pos.neg.spec.type.df$percent.pos<-round(pos.neg.spec.type.df$Positive/(pos.neg.spec.type.df$Negative + pos.neg.spec.type.df$Positive)*100,2)
print(pos.neg.spec.type.df)

pos.neg.spec.type.df$specimen_type<-rownames(pos.neg.spec.type.df)
#write.xlsx(pos.neg.spec.type.df,
           #file = "../../manuscripts/prolonged_positive_manuscript/JCM_submission/figures_and_tables/2021-09-03_percent_positive_by_specimen_type.xlsx")

print('Total unique patients tested:')
length(unique(covid.df$study.id))


print("Unique patients with at least 1 positive test at any time: ")
# positive at any time
covid.pos.pts<-as.character(covid.df$study.id)[covid.df$Value == "Positive" | covid.df$Value == "Detected" ]
print(length(unique(covid.pos.pts)))

print("Patients who were tested more than once: ")
patients.mult.tests<-names(table(covid.df$study.id)[table(covid.df$study.id) >1])
print(length(unique(patients.mult.tests)))


print("Patients who were tested more than once including at least 1 positive at any time: ")
print(length(unique(patients.mult.tests[patients.mult.tests %in% covid.pos.pts])))

print("Patients with >1 positive test:")
print(length(unique(mult.pos.df$study.id)))



print("Interval between first and last positive tests (days) for patients who tested positive more than once: ")
print(summary(as.numeric(repeat.test.df$first.last.pos.days[!duplicated(repeat.test.df$study.id)])))

#Total patients with first and last positive tests w/in 1 incubation period (14 days)
print("Interval between first and last positive tests (days) within 1 incubation period (14 days) for patients who tested positive more than once: ")
print(summary(as.numeric(repeat.test.df$first.last.pos.days[!duplicated(repeat.test.df$study.id) & repeat.test.df$first.last.pos.days <= 14])))
print("N patients: ")
print(sum(!duplicated(repeat.test.df$study.id) & repeat.test.df$first.last.pos.days <= 14))


print("Interval between first and last positive tests (days) greater than the 3rd quartile for first and last positive date (28 days) for patients who tested positive more than once: ")
print(summary(as.numeric(repeat.test.df$first.last.pos.days[!duplicated(repeat.test.df$study.id) & repeat.test.df$first.last.pos.days >28])))
print("N patients: ")
print(sum(!duplicated(repeat.test.df$study.id) & repeat.test.df$first.last.pos.days >28))


print("Total patients with positive negative positive pattern: ")
print(length(unique(repeat.test.df$study.id[repeat.test.df$pos.test.count>1 & repeat.test.df$first.last.pos.days>1 & repeat.test.df$pos.neg.pos>0])))


print("number of intermittent negative tests between positive tests")




# generate spreadsheet for study.idS for study
# send 4/20 2021
#write.csv(x = repeat.test.df$study.id[!duplicated(repeat.test.df$study.id) & repeat.test.df$first.last.pos.days >28],file #= "data/chart_review_data/2021-03-12_full_cohort/2021-04-20_prolonged_covid_positive_57_pts.csv")


```


### Supplemental table 1: Summary of RT-PCR tests performed during study
-table w/specimen type and % positivity and platforms testing occured on

```{r,supplemental_table_1,fig.width=10,fig.height=11}


pos.neg.platform.df<-as.data.frame(do.call(cbind, by(as.vector(covid.df$specimen.Type.clean), covid.df$platform, summary)))

pos.neg.spec.type.df<-as.data.frame(do.call(cbind, by(as.vector(covid.df$specimen.Type.clean), covid.df$pos.test, summary)))


# add % positive column
pos.neg.spec.type.df$percent.pos<-round(pos.neg.spec.type.df$Positive/(pos.neg.spec.type.df$Negative + pos.neg.spec.type.df$Positive)*100,2)


sup.tab.1<-cbind(pos.neg.platform.df, pos.neg.spec.type.df)

knitr::kable(sup.tab.1, caption = 'Supplemental table 1: summary of RT-PCR tests performed during study')

#print(sup.tab.1)
#write.csv(sup.tab.1, file="figures/2021-06-12_supplemental_1.csv")
```



### Supplemental table 2: types of patients tested on each platform
-not 100% sure we need this table, will clean it up and collapse rows if we want to present this data

```{r, supplemental_table_2}

order.dept.platform<-as.matrix(table(as.vector(covid.df$Ordering.Department),as.vector(covid.df$platform)))

order.dept.platform<-order.dept.platform[order(rowSums(order.dept.platform)),]


knitr::kable(order.dept.platform, caption = 'Supplemental table 2: location of patients tested on each platform')

#write.csv(order.dept.platform, file="figures/2021-06-12_supplemental_2.csv")

```


### Figure 1A: SARS-CoV2 testing patterns for patients who tested positive >1 time during the study

```{r,Figure_1A, fig.width=10,fig.height=11}

repeat.test.df$Collected<-as.Date(repeat.test.df$Collected)

result.colors<-c("blue","red")

ggplot(repeat.test.df, aes(y=reorder(study.id, first.pos.date), x=Collected, group=study.id)) + 
  geom_jitter( height = 0.1,width = 0.0,size = 1, aes(fill=pos.test), colour="black", pch=21)+
  xlab("Time (week) ") +
  ylab("Patients") + 
  theme_bw()+
   theme( panel.grid.major.y = element_line( size=.3, color="darkgrey" ),axis.text.x = element_blank()) +
  ggtitle("Patients with at least 2 positive tests")+
  labs(fill='RT-PCR Result')+
scale_x_date(breaks = seq(min(repeat.test.df$Collected), max(repeat.test.df$Collected), by = 7), labels=seq(min(repeat.test.df$Collected), max(repeat.test.df$Collected), by =7))+
  theme(axis.text.x = element_text(angle = 90), axis.text.y=element_blank())+
  scale_fill_manual(values=result.colors)+
 scale_shape_manual(values = c(21,22,23,24))+
  xlab("Time (week) ") +
  ylab("Patients")


```


### Figure 1B: Distribution of timing between first and last positive test for patients who tested positive >1 time during the study

```{r,figure_1B, fig.width=5,fig.height=5}


repeat.test.df$third.quartile<-repeat.test.df$first.last.pos.days > 28

ggplot(data=subset(repeat.test.df, !(duplicated(study.id))), aes(y=first.last.pos.days, x = factor(Testing.Lab))) +
  geom_violin(width=.7, adjust=2)+
  geom_boxplot(width=.3, color="black",outlier.size=-1)+
  geom_jitter(height = 0.0,width = 0.1,size = 1.75,pch=21, aes(fill=third.quartile))+
  scale_fill_manual(values=c("#999999", "cornflowerblue"),labels = c("less than or equal to 28 days", "greater than 28 days"))+
  labs(fill="Days between first and last positive test") +
  ggtitle("Days between first and last positive test")+
  scale_y_continuous(breaks = scales::pretty_breaks(n = 20))+
    xlab("Tests from patients with \n multiple positive tests") +
  ylab("Days between first and last positive test")+
  theme_bw()+
theme(axis.text.x = element_blank())


ggsave(path = "fig_jpegs_and_table/",filename = "2022-03-02_figure_1",
      width =6.87, height =9.06, device='tiff', dpi=300)

```


## Results section 2: summarize characteristics of patients with prolonged positivity
-number of tests among patients with prolonged positivity
-intermittent positivity  
-timing between negative and subsequent positive for intermittant positives  
-specimen source for intermittent positives  
-types of specimens tested ovet time  
-Association between Ct values and specimen type tested over time  


```{r, summarise_characteristics_pts_prolonged_pos}
# patients with first/last positive >28 days apart during the study  
prolonged.pos.pts<-repeat.test.df$study.id[!duplicated(repeat.test.df$study.id) & repeat.test.df$first.last.pos.days >28]

prolonged.pos.df<-repeat.test.df[repeat.test.df$study.id %in% prolonged.pos.pts, ]

print("Number of tests among patients with prolonged positivity >28 days: ")
print(dim(prolonged.pos.df)[1])

print("number of negative and positive tests: ")
print(table(prolonged.pos.df$pos.test))

print("distribution of positive tests per patient: ")
 summary(prolonged.pos.df$pos.test.count[!(duplicated(prolonged.pos.df$study.id))])

print("patients with negative tests between positive tests: ")
table(prolonged.pos.df$pos.neg.pos[!duplicated(prolonged.pos.df$study.id)])

print("Number of negative tests per patient between positive tests for patients with intermittent positives: ")
summary(prolonged.pos.df$pos.neg.pos[!duplicated(prolonged.pos.df$study.id)])

#df subset for patient with prolonged positivity
prolonged.pos.pos.tests.df<-prolonged.pos.df[prolonged.pos.df$pos.test == "Positive", ]

```

### compare percentages of intemittant positivity among non-prolonged and prolonged positive patients

```{r}


# non prolonged
(33-16)/(264-57)*100

# prolonged



# prop.test
print("prop test for intermittant positives among non-prolonged and prolonged positive patients")
prop.test(x = c(17, 16), n = c(207, 57))
```




#### examine missing Ct values: wrong platform assigned?
- NOT SURE WHY WE ARE MISSING THESE CT VALUES??
- INVESTIGATE IN EPIC?
```{r, examine_mising_ct_values}



for(sid in manual.ct.values.df$MPI){

# current values 
current.ceph.e<-manual.ct.values.df$Cepheid_E[manual.ct.values.df$MPI ==sid]
current.ceph.n2<-manual.ct.values.df$Cepheid_N2[manual.ct.values.df$MPI ==sid]
current.alinity<-manual.ct.values.df$Alinity[manual.ct.values.df$MPI ==sid]


if(is.na(prolonged.pos.df$Cepheid_E[prolonged.pos.df$SID == sid]) & !is.na(current.ceph.e)){
  prolonged.pos.df$Cepheid_E[prolonged.pos.df$SID == sid]<-as.numeric(current.ceph.e)
  }


if(is.na(prolonged.pos.df$Cepheid_N2[prolonged.pos.df$SID == sid]) & !is.na(current.ceph.n2)){
  prolonged.pos.df$Cepheid_N2[prolonged.pos.df$SID == sid]<-as.numeric(current.ceph.n2)
  }

if(is.na(prolonged.pos.df$Alinity[prolonged.pos.df$SID == sid]) & !is.na(current.alinity)){
  prolonged.pos.df$Alinity[prolonged.pos.df$SID == sid]<-as.numeric(current.alinity)
}
  
}

#df subset for patient with prolonged positivity
prolonged.pos.pos.tests.df<-prolonged.pos.df[prolonged.pos.df$pos.test == "Positive", ]

missing.ct.test<-prolonged.pos.pos.tests.df[rowSums(prolonged.pos.pos.tests.df[,c("M2000","Alinity","Cepheid_E","Cepheid_N2", "Cepheid_RdRp")],na.rm = T) <1 & prolonged.pos.pos.tests.df$platform != "ML-MANUAL",]

#write table to find the missing Ct values by hand from instrument
find.missing.Ct<-missing.ct.test[,c("Resulted","study.id","Specimen.ID","Method","MPI","platform","Alinity",                        "Cepheid_RdRp","Cepheid_E","Cepheid_N2")]

paste0("patients withing at least 1 test with missing CT values: ", length(unique(missing.ct.test$study.id)))
paste0("orders from patients with CT value missing: ", length(unique(missing.ct.test$MPI)))

paste0("Percent of Ct values we have from positive tests from prolonged positive patients: ",round(((table(prolonged.pos.df$pos.test)[2]-length(unique(missing.ct.test$MPI)))/table(prolonged.pos.df$pos.test)[2])*100,2),"%")


```



### Figure 2A: Ct value and specimen type over time for patients with prolonged positivity
-plot CT positive tests over time since first positive by specimen type and platform  
-remove negative test dates for plotting

```{r,figure_2A, fig.width=7,fig.height=5}

# days between first positive and current test
prolonged.pos.df$current.test.to.first.pos<-prolonged.pos.df$Collected-prolonged.pos.df$first.pos.date

CN.plot.df<-prolonged.pos.df[,c("study.id","MPI","current.test.to.first.pos","pos.test","M2000","Alinity","Cepheid_E","Cepheid_N2","specimen.Type.clean")]

DF1 <- melt(CN.plot.df, id.var=c("study.id","MPI","pos.test","current.test.to.first.pos","specimen.Type.clean"))

#dummy number for negative tests
DF1$value[DF1$pos.test == "Negative"]<-45 
DF1<-DF1[DF1$value !=45,]

# make negative tests at above the limit of detection

DF1$value[DF1$variable == "M2000" & DF1$value >0 & DF1$value < 45 & !(is.na(DF1$value))]<-DF1$value[DF1$variable == "M2000" & DF1$value >0 & DF1$value < 45 & !(is.na(DF1$value)) ] + 10



DF1$value<-round(DF1$value,2)



ggplot(data=subset(DF1, current.test.to.first.pos >0 & value != 0), aes(y=value, x = current.test.to.first.pos)) +
  geom_jitter(height = 0.0,width = 0.15,size = 1.5 , aes(fill=value, shape = specimen.Type.clean), color = "black")+
  scale_shape_manual(values = c(21,22,23,24))+
  scale_fill_gradient(name = "Ct value", 
                        low = "red", high = "blue")+
  labs(fill="Ct value", shape = "Specimen type") +
  ggtitle("Days since first positive test")+
  scale_y_continuous(breaks = scales::pretty_breaks(n = 20),trans = 'log10')+
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10))+
    xlab("Days") +
  ylab("Ct value")+
  stat_smooth()+
  theme_bw()
#+
  #theme_bw(axis.text.x = element_text(angle=90, size = 2))

prolonged.pos.ct.formattted<-DF1

```



## specimen type for negative and positive test for prolonged positivity based on days since first positive test

```{r,specimen_type_negative_positive_ct,  warning=FALSE,fig.width=8,fig.height=8}
CN.plot.df<-prolonged.pos.df[,c("study.id","MPI","current.test.to.first.pos","pos.test","M2000","Alinity","Cepheid_E","Cepheid_N2","specimen.Type.clean")]

DF1 <- melt(CN.plot.df, id.var=c("study.id","MPI","pos.test","current.test.to.first.pos","specimen.Type.clean"))
DF1$value<-round(DF1$value,2)


ggplot(data=subset(DF1, current.test.to.first.pos >=0 & value != 0), aes(y=value, x = current.test.to.first.pos, group=specimen.Type.clean)) +
  geom_jitter(height = 0.0,width = 0.15,size = 1.5 , aes(fill=value, shape = specimen.Type.clean), color = "black")+
  scale_shape_manual(values = c(21,22,23,24))+
  scale_fill_gradient(name = "Cycle Number", 
                        low = "red", high = "blue")+
  labs(fill="Ct value", shape = "Specimen type") +
  ggtitle("Days since first positive test")+
  scale_y_continuous(breaks = scales::pretty_breaks(n = 20),trans = 'log10')+
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10))+
    xlab("Days") +
  ylab("Ct value")+
  facet_wrap(vars(specimen.Type.clean))+
  stat_smooth()+
  theme(axis.text.x = element_text(angle=90, size = 8),strip.text = element_text(size = 12,angle=90))



```



## Supplemental figure 1: Days since first positive test for patients WITHOUT prolonged positivity
-note scales of x axis are different 

```{r,supplemental_figure_1, fig.width=7,fig.height=5}

# days between first positive and current test
not.prolonged.pos.df<-repeat.test.df[!(repeat.test.df$study.id %in% prolonged.pos.pts), ]

not.prolonged.pos.df$current.test.to.first.pos<-not.prolonged.pos.df$Collected-not.prolonged.pos.df$first.pos.date

CN.plot.df<-not.prolonged.pos.df[,c("study.id","MPI","current.test.to.first.pos","pos.test","M2000","Alinity","Cepheid_E","Cepheid_N2","specimen.Type.clean")]

DF1 <- melt(CN.plot.df, id.var=c("study.id","MPI","pos.test","current.test.to.first.pos","specimen.Type.clean"))

#dummy number for negative tests
DF1$value[DF1$pos.test == "Negative"]<-45 


# make negative tests at above the limit of detection

DF1$value[DF1$variable == "M2000" & DF1$value >0 & DF1$value < 45 & !(is.na(DF1$value))]<-DF1$value[DF1$variable == "M2000" & DF1$value >0 & DF1$value < 45 & !(is.na(DF1$value)) ] + 10

DF1<-DF1[DF1$value !=45,]

DF1$value<-round(DF1$value,2)


ggplot(data=subset(DF1, current.test.to.first.pos >0 & value != 0), aes(y=value, x = current.test.to.first.pos)) +
  geom_jitter(height = 0.0,width = 0.15,size = 1.5 , aes(fill=value, shape = specimen.Type.clean), color = "black")+
  scale_shape_manual(values = c(21,22,23,24,25,19,8,12))+
  scale_fill_gradient(name = "Cycle Number", 
                        low = "red", high = "blue")+
  labs(fill="Ct value", shape = "Specimen type") +
  ggtitle("Days since first positive test")+
  scale_y_continuous(breaks = scales::pretty_breaks(n = 20),trans='log10')+
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10))+
    xlab("Days") +
  ylab("Ct value")+
  stat_smooth()
#+
  #theme_bw(axis.text.x = element_text(angle=90, size = 2))



#bins for weeks since first positive test
DF1<-DF1[DF1$current.test.to.first.pos>0,]
DF1$week.bin<-ceiling(DF1$current.test.to.first.pos/7)
DF1<-DF1[DF1$pos.test=="Positive",]


#DF1$week.bin[DF1$week.bin ==5]<-5

day.labs<-c("1-7","8-14","15-21","22-28")

DF1$specimen.Type.clean[!(DF1$specimen.Type.clean %in% 'Nasopharyngeal Swab') & !is.na(DF1$specimen.Type.clean)]<-"Other"

ggplot(data=subset(DF1, !(is.na(week.bin))), aes(x=1, y = value)) +
  geom_violin(width=.7, adjust=2)+
  geom_boxplot(width=.1, color="black",outlier.size=-1)+
  geom_jitter(height = 0.0,width = 0.24,size = 1.5,aes(shape = specimen.Type.clean),color="black")+
  scale_shape_manual(values = c(20,21,22,23,24))+
  labs(shape = "Specimen type") +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 15))+
    xlab("1-28") +
  ylab("Cycle threshold")+
  theme_bw()+
  scale_x_discrete(labels= day.labs)+
  ggtitle("Patients without prolonged positivity")

ggsave(path = "fig_jpegs_and_table/",filename = "2022-03-02_figure_2A",
        width =6.87, height =9.06, device='tiff', dpi=300)

ct.formatted.pts.no.prolonged.pos<-DF1

```




## Figure 2B: ct trend by specimen type over time linked to each patient for patients with intermittant positive tests

for individual patients need to plot unknown values from LDT tests or it is missleading.
-__NOTE: patients may have <28 days represented because we do not have Ct values for LDT and other missing Cts__


```{r,figure_2B, fig.width=8,fig.height=8}

# days between first positive and current test
prolonged.pos.df$current.test.to.first.pos<-prolonged.pos.df$Collected-prolonged.pos.df$first.pos.date

CN.plot.df<-prolonged.pos.df[,c("study.id","MPI","current.test.to.first.pos","pos.test","M2000","Alinity","Cepheid_E","Cepheid_N2","specimen.Type.clean")]

DF1 <- melt(CN.plot.df, id.var=c("study.id","MPI","pos.test","current.test.to.first.pos","specimen.Type.clean"))

#dummy number for negative tests
DF1$value[DF1$pos.test == "Negative"]<-45 
#DF1<-DF1[DF1$value !=45,]

# make negative tests at above the limit of detection

DF1$value[DF1$variable == "M2000" & DF1$value >0 & DF1$value < 45 & !(is.na(DF1$value))]<-DF1$value[DF1$variable == "M2000" & DF1$value >0 & DF1$value < 45 & !(is.na(DF1$value)) ] + 10



DF1$value<-round(DF1$value,2)


prolonged.int.pos.pt<-unique(prolonged.pos.df$study.id[!duplicated(prolonged.pos.df$study.id) & prolonged.pos.df$pos.neg.pos >0])


DF1$study.id.factor<-as.factor(DF1$study.id)
levels(DF1$study.id.factor)<-1:57

ggplot(data=subset(DF1, current.test.to.first.pos >=0 & value != 0 & study.id %in% prolonged.int.pos.pt), aes(y=value, x = current.test.to.first.pos, group=study.id)) +
  geom_jitter(height = 0.0,width = 0.15,size = 1.5 , aes(fill=value, shape = specimen.Type.clean), color = "black")+
  scale_shape_manual(values = c(21,22,23,24))+
  scale_fill_gradient(name = "Cycle Number", 
                        low = "red", high = "blue")+
  labs(fill="Ct value", shape = "Specimen type") +
  ggtitle("Days since first positive test: prolonged positive patients with intermittent positive tests")+
scale_y_continuous(trans="log10")+
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10))+
    xlab("Days from first positive test") +
  ylab("Ct value")+
  #stat_smooth()+
  geom_line(size = .5)+
  facet_wrap(~study.id.factor)+
   theme(axis.text.x=element_text(angle=90,hjust=1))+
geom_hline(yintercept=45, linetype="dashed", 
                color = "red", size=1)


```

 
 
### Summarize time between positive and negative tests for intermittent positive patients with prolonged positivity
-first positive and intermittent negatives before last positive
-last positive and intermittent negatives before last positive
-intermittent positives and negatives between first and last positive


```{r,time_between_positive_tests_intermittent, warning=FALSE}

#time between first positive and intermittent negative tests before last positive
int.neg.pos.dates<-sapply(prolonged.int.pos.pt, FUN=function(pt){
  
  pt.df<-prolonged.pos.df[prolonged.pos.df$study.id==pt,]
  
  all.neg.dates<-as.Date(pt.df$Received)[pt.df$pos.test == "Negative" & pt.df$Collected > pt.df$first.pos.date & pt.df$Collected < pt.df$last.pos.date]

  unique(all.neg.dates-pt.df$first.pos.date)
  
})
names(int.neg.pos.dates)<-prolonged.int.pos.pt

summary(unlist(int.neg.pos.dates))



#time between last positive and intermittent negative tests before last positive
int.neg.pos.dates<-sapply(prolonged.int.pos.pt, FUN=function(pt){
  
  pt.df<-prolonged.pos.df[prolonged.pos.df$study.id==pt,]
  
  all.neg.dates<-as.Date(pt.df$Received)[pt.df$pos.test == "Negative" & pt.df$Collected > pt.df$first.pos.date & pt.df$Collected < pt.df$last.pos.date]

  unique(all.neg.dates-pt.df$last.pos.date)
  
})
names(int.neg.pos.dates)<-prolonged.int.pos.pt

print("Time between last positive anf intermittent negative test before last positive: ")
summary(unlist(int.neg.pos.dates))



#time between intermittent positive and intermittent negative tests before last positive
int.neg.pos.neg.dates<-sapply(prolonged.int.pos.pt, FUN=function(pt){
  
  pt.df<-prolonged.pos.df[prolonged.pos.df$study.id==pt,]
  
  #positives before first and last positive
  all.pos.dates<-as.Date(pt.df$Received)[pt.df$pos.test == "Positive" & pt.df$Collected > pt.df$first.pos.date & pt.df$Collected < pt.df$last.pos.date]
#negatives before first and last positive
    all.neg.dates<-as.Date(pt.df$Received)[pt.df$pos.test == "Negative" & pt.df$Collected > pt.df$first.pos.date & pt.df$Collected < pt.df$last.pos.date]
  
  unique(all.pos.dates-all.neg.dates)
  
})
names(int.neg.pos.neg.dates)<-prolonged.int.pos.pt

print("Time between intermittent positive and intermittend negative tests before last positive: ")
summary(unlist(int.neg.pos.neg.dates))



```

### Summarize specimen type for intermittent positive test
 
```{r, specimen_type_intermittent_positive_tests, warning=FALSE}

int.neg.spec.type<-sapply(prolonged.int.pos.pt, FUN=function(pt){
  
  pt.df<-prolonged.pos.df[prolonged.pos.df$study.id==pt,]
  

#negatives before first and last positive
    all.neg.dates<-as.Date(pt.df$Received)[pt.df$pos.test == "Negative" & pt.df$Collected > pt.df$first.pos.date & pt.df$Collected < pt.df$last.pos.date]
    
  
  as.vector(pt.df$specimen.Type.clean)[pt.df$Collected == all.neg.dates]
  
})
names(int.neg.spec.type)<-prolonged.int.pos.pt

print("Specimen types for negatives between first and last positive: ")
table(unlist(int.neg.spec.type))


int.pos.spec.type<-sapply(prolonged.int.pos.pt, FUN=function(pt){
  
  pt.df<-prolonged.pos.df[prolonged.pos.df$study.id==pt,]
  

#positives between  before first and last positive
  
  int.pos.spec.type<-sapply(prolonged.int.pos.pt, FUN=function(pt){
  
  pt.df<-prolonged.pos.df[prolonged.pos.df$study.id==pt,]
  

  
    all.pos.dates<-as.Date(pt.df$Received)[pt.df$pos.test == "Positive" & pt.df$Collected > pt.df$first.pos.date & pt.df$Collected < pt.df$last.pos.date]
    
  
  as.vector(pt.df$specimen.Type.clean)[pt.df$Collected == all.pos.dates]
  
})
})


names(int.pos.spec.type)<-prolonged.int.pos.pt

print("Specimen types for positives between first and last positive: ")
table(unlist(int.pos.spec.type))

```
 

 
### Check that full data set is included in clinical data, should be 57 patients with prolonged positivity
-medication data should have 50 patients according to CDWH


```{r, check_clinical_data}
print(paste0("Unique patients represented in clinical dataset: ", length(unique(clinical.df$study.id))))

print(paste0("Prolonged positive patients missing from clinical chart review dataset : ", length(setdiff(unique(prolonged.pos.pts), unique(clinical.df$study.id)))))

print(paste0("Unique patients represented in meds dataset: ",  length(unique(meds.df$study.id))))

```

 
 ### compare basic demographics for patients with and without prolonged positivity  
 -age  
 -gender  
 
```{r, compare_demographics_with_without_prolonged_positivity}
pts.not.prolonged<-setdiff(covid.df$study.id, prolonged.pos.pts)

other.ages<-as.numeric(unlist(regmatches(covid.df$Age, gregexpr("[[:digit:]]+", covid.df$Age))))

# (minimum, lower-hinge, median, upper-hinge, maximum)

# other patients
fivenum(other.ages[as.vector(covid.df$study.id) %in% pts.not.prolonged])

#prolonged positive patients
fivenum(other.ages[as.vector(covid.df$study.id) %in% prolonged.pos.pts])

print("Gender distribution for not prolonged positive patients:  ")
table(covid.df$Gender[as.vector(covid.df$study.id) %in% pts.not.prolonged])

print("Gender distribution for prolonged positive patients:  ")
table(covid.df$Gender[as.vector(covid.df$study.id) %in% prolonged.pos.pts])

```
 
 
 
### Check and subset medication data for drugs of interest during study timeframe  
-patients without meds in spreadsheet  
-patients with meds in spreadsheet  
-subset medications for patients of interest based on meds that were looked at by Suba and Billy for the manual chart review  
-subset for study timeframe  

__chart review drugs: __  
Tacrolimus (Prograf)  
mycophenolate (Cellcept)  
hydroxycloroquine,  
prednisone,  
tocilizumab,  
methotrexate,  
Mercaptopurine,  
Fluticasone,  
Lopinavir / Ritonavir,  
Canakinumab,  
Immunoglobulin,  
Cyclophosphamide  

__Drugs added by me:__    
-remdesivir  
-Dexamethasone  
 
```{r,format_drugs}


drugs<-c("hydroxychloroquine",
         "prednisone",
         "Tacrolimus",
         "mycophenolate",
         "tocilizumab",
         "Plaquenil",
         "methotrexate",
         "Mercaptopurine",
         "Fluticasone",
         "Lopinavir",
         "Ritonavir",
         "Canakinumab",
         "Immunoglobulin",
         "Cyclophosphamide",
         "remdesivir",
         "Dexamethasone")


med.row<-sapply(drugs, FUN=function(med){
  
  #return logical vector
  grepl(med,meds.df$NAME,ignore.case=T)

})


# rows containing drugs of interest
subset.meds.df<-meds.df[rowSums(med.row,na.rm = T)>0,]



# subset for meds exposures during the study
#format dates
subset.meds.df$START_DATE<-as.Date.character(subset.meds.df$START_DATE,format = "%m/%d/%y" )
subset.meds.df$END_DATE<-as.Date.character(subset.meds.df$END_DATE,format = "%m/%d/%y" )


# subset by drugs exposed between min and max collected date for prolonged positive patients
#logical matrix indicating drug exposure during timeframe
subset.med.row<-med.row[rowSums(med.row,na.rm = T)>0,]


subset.med.row<-subset.med.row[subset.meds.df$START_DATE<=max(prolonged.pos.pos.tests.df$Collected) & subset.meds.df$START_DATE>=min(prolonged.pos.pos.tests.df$Collected),]
subset.med.row<-subset.med.row[!is.na(subset.med.row[,1]),]

#subset full meds.df for timeframe
subset.meds.df<-subset.meds.df[subset.meds.df$START_DATE<=max(prolonged.pos.pos.tests.df$Collected) & subset.meds.df$START_DATE>=min(prolonged.pos.pos.tests.df$Collected), ]


# add column for clean drug names
subset.meds.df$clean.drug.name<-sub(" .*",'',subset.meds.df$NAME)

#remove NAs?
subset.meds.df<-subset.meds.df[!(is.na(subset.meds.df$study.id)),]


#subset.meds.df<-subset.meds.df[!(is.na(subset.meds.df)) & ,]

# check number of doses for each drug
drug.doses<-colSums(med.row)

#unique patients exposed to drugs during the study

drug.patients<-sapply(names(drug.doses),FUN=function(drug){
  
  unique(subset.meds.df$study.id[grepl(drug,subset.meds.df$NAME,ignore.case = T)])
  
})

knitr::kable(drug.doses, caption = 'Unique drug doses')
knitr::kable(unlist(lapply(drug.patients,length)),col.names = 'Drugs', caption = 'Patients with at least 1 dose of drug')

drug.doses<-as.data.frame(drug.doses)
drug.doses$drugs<-rownames(drug.doses)


#write.xlsx(x=drug.doses,file ="../../manuscripts/prolonged_positive_manuscript/JCM_submission/figures_and_tables/2021-09-03_medications_patients_exposed_during_study.xlsx")

```


### Make meds table
```{r}


# add study.id to logical matrix indicating date
subset.med.row<-cbind.data.frame(subset.med.row,
      study.id=subset.meds.df$study.id,
      start.date=subset.meds.df$START_DATE,
      end.date=subset.meds.df$END_DATE,
      frequency.name=subset.meds.df$FREQ_NAME)


```


## Plot medications over covid testing dates  

```{r,fig.height=15, fig.width=12}
# 
# 
# subset.meds.df$first.pos.date<-sapply(subset.meds.df$study.id, FUN=function(pt){
#   prolonged.pos.df$first.pos.date[prolonged.pos.df$study.id == pt][1]
# })
# 
# 
# subset.meds.df$pos.neg.pos<-sapply(subset.meds.df$study.id, FUN=function(pt){
#   prolonged.pos.df$pos.neg.pos[prolonged.pos.df$study.id == pt][1]
# })
# 
# subset.meds.df$Gender<-sapply(subset.meds.df$study.id, FUN=function(pt){
#   prolonged.pos.df$Gender[prolonged.pos.df$study.id == pt][1]
# })
# 
# 
# subset.meds.df$pharm.class.plotting<-subset.meds.df$PHARM_CLASS_NAME
# subset.meds.df$pharm.class.plotting[subset.meds.df$pharm.class.plotting=="NULL"]<-subset.meds.df$clean.drug.name[subset.meds.df$pharm.class.plotting=="NULL"]
# 
# drug.class.plot<-as.data.frame(table(subset.meds.df$clean.drug.name,subset.meds.df$pharm.class.plotting))
# drug.class.plot<-drug.class.plot[drug.class.plot$Freq>0, ]
# #drugs are given different classes, clean this up
# drug.class.plot[order(drug.class.plot$Var1),]
# 
# subset.meds.df$pharm.class.plotting[subset.meds.df$pharm.class.plotting == "GLUCOCORTICOIDS, ORALLY INHALED"]<-"GLUCOCORTICOIDS"
# 
# subset.meds.df$pharm.class.plotting[subset.meds.df$pharm.class.plotting == "TOCILIZUMAB"]<-"IMMUNOSUPPRESSIVES"
#                                     
# subset.meds.df$study.id<-subset.meds.df$study.id                                    
# 
# slow.ct.pts<-rownames(pt.factors.df)[pt.factors.df$Last_Positive_Ct<35 & pt.factors.df$first_last_pos_day>=40]
# slow.ct.pts<-slow.ct.pts[!is.na(slow.ct.pts)]
# 
ggplot(subset(prolonged.pos.df,study.id %in% slow.ct.pts), aes(y=reorder(study.id, first.pos.date), x=Collected, group=study.id)) +
  geom_jitter( height = 0.1,width = 0.0,size = 2, aes(fill=pos.test), colour="black", pch=21)+
  geom_text_repel(aes(label =Alinity ,angle=85),size=2,max.overlaps = 100,force=1, point.padding=unit(2,'lines'),
                direction = 'y',
                nudge_y = 0.2,
                segment.size=0.2)+
  geom_text_repel(aes(label =Cepheid_RdRp ,angle=85),size=3,max.overlaps = 100,force=1, point.padding=unit(2,'lines'),
                direction = 'y',
                nudge_y = 0.2,
                segment.size=0.2)+
  geom_text_repel(aes(label =M2000 ,angle=85),size=3,max.overlaps = 100,force=1, point.padding=unit(2,'lines'),
                direction = 'y',
                nudge_y = 0.2,
                segment.size=0.2)+
  geom_text_repel(aes(label =Cepheid_N2 ,angle=85),size=3,max.overlaps = 100,force=1, point.padding=unit(2,'lines'),
                direction = 'y',
                nudge_y = 0.2,
                segment.size=0.2)+
  geom_text_repel(aes(label =Cepheid_E ,angle=85),size=3,max.overlaps = 100,force=1, point.padding=unit(2,'lines'),
                direction = 'y',
                nudge_y = 0.2,
                segment.size=0.2)+

 xlab("Time (week) ") +
ylab("Patients") +
  labs(fill = "SCV-2 test result") +
theme_bw()+
theme( panel.grid.major.y = element_line( size=.3, color="darkgrey" ),axis.text.x = element_blank()) +
scale_x_date(breaks = seq(min(repeat.test.df$Collected), max(repeat.test.df$Collected), by = 7), labels=seq(min(repeat.test.df$Collected), max(repeat.test.df$Collected), by =7))+
theme(axis.text.x = element_text(angle = 90), axis.text.y=element_blank())+
scale_fill_manual(values=result.colors)
#  
# 

```



### Clean up admission data
-make naming of locations for disposition consistant

```{r}

admission.df$Discharge_disposition_clean<-admission.df$Discharge_disposition
admission.df$Discharge_disposition_clean[is.na(admission.df$Discharge_disposition) | admission.df$Discharge_disposition == "N/A"]<-NA
admission.df$Discharge_disposition_clean[admission.df$Discharge_disposition_clean %in% c("Skilled nursing facility (SNF)","SNF","Skilled nursing facility", " Skilled nursing facility")]<-"SNF"

admission.df$Discharge_disposition_clean[grepl("hospice",admission.df$Discharge_disposition_clean, ignore.case = T)]<-"Hospice"


admission.df$Discharge_disposition_clean[grepl("assist",admission.df$Discharge_disposition_clean, ignore.case = T)]<-"Assisted Living"

admission.df$Discharge_disposition_clean[grepl("home",admission.df$Discharge_disposition_clean, ignore.case = T)]<-"Home"
admission.df$Discharge_disposition_clean[grepl("Against",admission.df$Discharge_disposition_clean, ignore.case = T)]<-"Home"


admission.df$Discharge_disposition_clean[grepl("deceased",admission.df$Discharge_disposition_clean, ignore.case = T)]<-"Deceased"

admission.df$Discharge_disposition_clean[admission.df$Discharge_disposition_clean %in% "Med Hosp H (MDH)"]<-"Hospital"


admission.df$Discharge_disposition_clean<-as.vector(admission.df$Discharge_disposition_clean)

admission.df$Discharge_disposition_clean[grepl("acute",admission.df$Discharge_disposition_clean, ignore.case = T)]<-"Rehab"
# patient still in hospital
admission.df$Discharge_disposition_clean[admission.df$Notes == "Still admitted (1/8/2021)"]<-"Hospital"

admission.df$Discharge_disposition_clean[grepl("hospice",admission.df$Discharge_disposition, ignore.case = T)]<-"Hospice"
admission.df$Discharge_disposition_clean[admission.df$Discharge_disposition_clean=='']<-NA

#make the person who is still in the hospital have a discharge date of Oct 17 2020 vs NA
admission.df$Discharge_date[admission.df$Notes == "Still admitted (1/8/2021)"]<-"10/17/20"

# if missing discharge info but admission date 
admission.df$Discharge_disposition_clean[(is.na(admission.df$Discharge_date)) & is.na(admission.df$Discharge_disposition_clean)]<-"Never Admitted"

```



##Table N: summary of characteristics of patients with prolonged positivity: Patient demographic, hospitalization, other infection and treatment characteristics 

__IN PROGRESS__ 
-still adding variables.  

Distribution of symptomatic at first and subsequent tests  
Distribution of reasons for secondary testing at later timepoints   
Summary of clinical features:  
- hospitalizations  
-immunocompromised  
-treated w/antiviral drugs, drugs listed above  
-other infections during hospitalization during covid  
-deceased after covid   
-discharge disposition: last discharge
-cycle threshold at first positive  
-cycle threshold at last positive  
-intermittent negative between positives  

??? hospice?!

```{r}
# clinical table N patients x variables of interest

admission.df$Admission_date<-as.vector(admission.df$Admission_date)
admission.df$Admission_date[admission.df$Admission_date=='']<-NA
admission.df$Discharge_date<-as.vector(admission.df$Discharge_date)
admission.df$Discharge_date[admission.df$Discharge_date=='']<-NA
admission.df$LOS<-as.Date(admission.df$Discharge_date,format ="%m/%d/%y" )-as.Date(admission.df$Admission_date,format="%m/%d/%y")


# fix BMI character!
#clinical.df.save<-clinical.df

clinical.df$BMI<-as.vector(clinical.df$BMI)
clinical.df$BMI<-sub(" kg/m²",'',clinical.df$BMI)
clinical.df$BMI[clinical.df$BMI=="NONE"]<-NA
clinical.df$BMI<-as.numeric(clinical.df$BMI)


pt.factors.df<-t(sapply(unique(prolonged.pos.pts),FUN=function(pt){
  
  # age
  pt.age<-clinical.df$Age[clinical.df$study.id==pt][1]
  
  #sex
  pt.sex<-as.vector(prolonged.pos.df$Gender)[prolonged.pos.df$study.id==pt][1]
  
  #BMI
  pt.bmi<-clinical.df$BMI[clinical.df$study.id==pt][1]
  
  #admitted
  pt.admit.all<-admission.df$Admission_date[admission.df$study.id==pt & !is.na(admission.df$Admission_date)]
  
  if(sum(!is.na(pt.admit.all))>0){
      pt.admit<-"Admitted"
  }else{
    pt.admit<-"Never admitted"
  } #end else

  #admission LOS
  pt.los<-sum(admission.df$LOS[admission.df$study.id==pt],na.rm = T)

  # days between first and last positive
  pt.first.last.pos.day<-prolonged.pos.df$first.last.pos.days[prolonged.pos.df$study.id==pt][1]
  
  
  #first positive Ct value
pt.first.pos.Ct<-max(prolonged.pos.ct.formattted$value[prolonged.pos.ct.formattted$study.id==pt & prolonged.pos.ct.formattted$current.test.to.first.pos==0],na.rm = T)
  
  
  #last test Ct value
  last.pos.day.pt<-as.numeric(prolonged.pos.df$first.last.pos.days[prolonged.pos.df$study.id==pt][1] )
  

# ct value on last positive day for the patient
  pt.last.pos.Ct<-as.vector(max(prolonged.pos.ct.formattted$value[prolonged.pos.ct.formattted$study.id==pt & prolonged.pos.ct.formattted$current.test.to.first.pos==last.pos.day.pt],na.rm = T))
  

  #change in Ct value
 
  #drugs
  
  
  
  #discharge disposition at last discharge
  if(pt.admit=="Admitted"){
    
  pt.final.disch<-max(admission.df$Discharge_date[admission.df$study.id==pt],na.rm = T)
  
  final.discharge.dispo<-sort(admission.df$Discharge_disposition_clean[admission.df$study.id==pt & admission.df$Discharge_date==pt.final.disch])
  
  if(length(final.discharge.dispo)==0){
    final.discharge.dispo<-NA
  }

  } else{

    final.discharge.dispo<-"Never Admitted"
  }


  # length of stay in hospital

  
  #other infections
  
  
  out.vec<-c(pt.age,
             pt.bmi, 
             pt.sex,
             pt.admit,
             pt.los,
            final.discharge.dispo,
             pt.first.pos.Ct,
             pt.last.pos.Ct,
             pt.first.last.pos.day)
  names(out.vec)<-c("Age",
                    "BMI",
                    "Sex",
                    "Admitted",
                    "Length_of_stay",
                  "Discharge_disposition",
                    "First_Positive_Ct",
                    "Last_Positive_Ct",
                    "first_last_pos_day")
  return(out.vec)
  
}))

rownames(pt.factors.df)<-unique(prolonged.pos.pts)
pt.factors.df<-as.data.frame(pt.factors.df)



## Format table 1


pt.factors.df$first_last_pos_day<-as.numeric(as.vector(pt.factors.df$first_last_pos_day))
label(pt.factors.df$First_Positive_Ct)<- "Days between first and last positive"
pt.factors.df$first_last_pos_day<-as.numeric(as.vector(pt.factors.df$first_last_pos_day))
units(pt.factors.df$first_last_pos_day)<- "days"
label(pt.factors.df$first_last_pos_day) <- "Time between first and last positive"



pt.factors.df$First_Positive_Ct<-as.numeric(as.vector(pt.factors.df$First_Positive_Ct))
pt.factors.df$First_Positive_Ct[is.infinite(pt.factors.df$First_Positive_C)]<-NA
label(pt.factors.df$First_Positive_Ct)<- "Cycle threshold first positive"

pt.factors.df$Last_Positive_Ct<-as.numeric(as.vector(pt.factors.df$Last_Positive_Ct))
pt.factors.df$Last_Positive_Ct[is.infinite(pt.factors.df$Last_Positive_C)]<-NA
label(pt.factors.df$Last_Positive_Ct)<- "Cycle threshold last positive"



pt.factors.df$Age<-as.numeric(as.vector(pt.factors.df$Age))
units(pt.factors.df$Age)<- "years"
label(pt.factors.df$Age) <- "Age"

pt.factors.df$BMI<-as.numeric(as.vector(pt.factors.df$BMI))
units(pt.factors.df$BMI)<- "kg/m^2"
label(pt.factors.df$BMI) <- "BMI"





my.render.cont <- function(x) {
    with(stats.apply.rounding(stats.default(x), digits=2), c("",
        "Mean (SD)"=sprintf("%s (&plusmn; %s)", MEAN, SD)))
}
my.render.cat <- function(x) {
    c("", sapply(stats.default(x), function(y) with(y,
        sprintf("%d (%0.0f %%)", FREQ, PCT))))
}


pdf(file = 'summary_table.pdf')
table1(~ Sex + BMI + Age + First_Positive_Ct + Last_Positive_Ct + first_last_pos_day + Admitted + Discharge_disposition, data=pt.factors.df,
       render.continuous=my.render.cont, render.categorical=my.render.cat)
dev.off()
```



### Figure 2: Ct range by first through Nth test violin plot
-try binning by days N=7

```{r,Figure_1_Ct_range_by_day_since_first_positive_violin_plot}

CN.plot.df<-prolonged.pos.df[,c("study.id","MPI","current.test.to.first.pos","pos.test","M2000","Alinity","Cepheid_E","Cepheid_N2","specimen.Type.clean")]

DF1 <- melt(CN.plot.df, id.var=c("study.id","MPI","pos.test","current.test.to.first.pos","specimen.Type.clean"))
DF1$value<-round(DF1$value,2)
#bins for weeks since first positive test
DF1<-DF1[DF1$current.test.to.first.pos>0,]
DF1$week.bin<-ceiling(DF1$current.test.to.first.pos/28)
DF1<-DF1[DF1$pos.test=="Positive",]

# add 10 cycles for M2000
DF1$value[DF1$variable == "M2000"]<-DF1$value[DF1$variable == "M2000"]+10




DF1$week.bin[DF1$week.bin ==6]<-5

day.labs<-c("1-28","29-56","57-85","86-112",">=113")

DF1$specimen.Type.clean[!(DF1$specimen.Type.clean %in% 'Nasopharyngeal Swab') & !is.na(DF1$specimen.Type.clean)]<-"Other"


ggplot(data=DF1, aes(x=factor(round(week.bin)), y = value)) +
  geom_violin(width=.7, adjust=2)+
  geom_boxplot(width=.1, color="black",outlier.size=-1)+
  geom_jitter(height = 0.0,width = 0.24,size = 1.5,aes(shape = specimen.Type.clean),color="black")+
  scale_shape_manual(values = c(20,21,22,23,24))+
  labs(shape = "Specimen type") +
  scale_y_continuous(breaks = seq(14,46,2))+
    xlab("Days from first positive test") +
  ylab("Cycle threshold")+
  theme_bw()+
  scale_x_discrete(labels= day.labs)+
  ggtitle("Patients with prolonged positivity")

ggsave(path = "fig_jpegs_and_table/",filename = "2022-03-02_figure_2A",
       width =6.87, height =9.06, device='tiff', dpi=300)
```


### Summarize Ct values for patients without prolonged positivty
```{r}

# days between first positive and current test
not.prolonged.pos.df<-repeat.test.df[!(repeat.test.df$study.id %in% prolonged.pos.pts), ]

not.prolonged.pos.df$current.test.to.first.pos<-not.prolonged.pos.df$Collected-not.prolonged.pos.df$first.pos.date

CN.plot.df<-not.prolonged.pos.df[,c("study.id","MPI","current.test.to.first.pos","pos.test","M2000","Alinity","Cepheid_E","Cepheid_N2","specimen.Type.clean","first.last.pos.days")]

DF1 <- melt(CN.plot.df, id.var=c("study.id","MPI","pos.test","current.test.to.first.pos","specimen.Type.clean","first.last.pos.days"))



first.pos.ct<-sapply(unique(DF1$study.id),FUN=function(pt){
  
  max(DF1$value[DF1$study.id==pt & DF1$current.test.to.first.pos==0],na.rm = T)
  
})


 last.pos.ct<-sapply(unique(DF1$study.id),FUN=function(pt){ 
   
    last.pos.day.pt<-as.numeric(DF1$first.last.pos.days[DF1$study.id==pt][1])
  
  
    # ct value on last positive day for the patient
  pt.last.pos.Ct<-as.vector(max(DF1$value[DF1$study.id==pt & DF1$current.test.to.first.pos==last.pos.day.pt],na.rm = T))

    
   })
  #last test Ct value
 
 
  last.pos.day.pt<-sapply(unique(DF1$study.id),FUN=function(pt){ 
   
    last.pos.day.pt<-as.numeric(DF1$first.last.pos.days[DF1$study.id==pt][1])
  })
 

not.prolonged.pos.ct<-as.data.frame(cbind(unique(DF1$study.id), first.pos.ct, last.pos.ct,last.pos.day.pt))
not.prolonged.pos.ct$delta.ct<-not.prolonged.pos.ct$last.pos.ct-not.prolonged.pos.ct$first.pos.ct


summary(not.prolonged.pos.ct$delta.ct[is.finite(not.prolonged.pos.ct$delta.ct)])

# patients with lower ct/more virus at last test vs first.

length(unique(not.prolonged.pos.ct$V1)[not.prolonged.pos.ct$delta.ct<0 &  is.finite(not.prolonged.pos.ct$delta.ct)])

not.prolonged.pos.ct$last.pos.day.pt[not.prolonged.pos.ct$delta.ct<0 &  is.finite(not.prolonged.pos.ct$delta.ct)]

summary(not.prolonged.pos.ct$last.pos.day.pt[not.prolonged.pos.ct$delta.ct<0 &  is.finite(not.prolonged.pos.ct$delta.ct)])

```





### For CVS abstract

```{r}

# terminal positive ct for patients with prolonged positivity
sum(pt.factors.df$Last_Positive_Ct>35 ,na.rm = T)
#[1] 33
sum(pt.factors.df$Last_Positive_Ct>35 ,na.rm = T)/(length(pt.factors.df$Last_Positive_Ct[!is.na(pt.factors.df$Last_Positive_Ct)]))


# obese count
length(pt.factors.df$BMI[pt.factors.df$BMI>=30 & !(is.na(pt.factors.df$BMI))])


# admitted
pt.factors.df$Admitted


# discharge to hospice or deceased

length(unique(admission.df$study.id[admission.df$Discharge_disposition_clean %in% c("Assisted Living","SNF")]))



```


## Figure 4 testing with Ct over time


```{r,fig.height=15, fig.width=12}


subset.meds.df$first.pos.date<-sapply(subset.meds.df$study.id, FUN=function(pt){
  prolonged.pos.df$first.pos.date[prolonged.pos.df$study.id == pt][1]
})


subset.meds.df$pos.neg.pos<-sapply(subset.meds.df$study.id, FUN=function(pt){
  prolonged.pos.df$pos.neg.pos[prolonged.pos.df$study.id == pt][1]
})

subset.meds.df$Gender<-sapply(subset.meds.df$study.id, FUN=function(pt){
  prolonged.pos.df$Gender[prolonged.pos.df$study.id == pt][1]
})


subset.meds.df$pharm.class.plotting<-subset.meds.df$PHARM_CLASS_NAME
subset.meds.df$pharm.class.plotting[subset.meds.df$pharm.class.plotting=="NULL"]<-subset.meds.df$clean.drug.name[subset.meds.df$pharm.class.plotting=="NULL"]

drug.class.plot<-as.data.frame(table(subset.meds.df$clean.drug.name,subset.meds.df$pharm.class.plotting))
drug.class.plot<-drug.class.plot[drug.class.plot$Freq>0, ]
#drugs are given different classes, clean this up
drug.class.plot[order(drug.class.plot$Var1),]

subset.meds.df$pharm.class.plotting[subset.meds.df$pharm.class.plotting == "GLUCOCORTICOIDS, ORALLY INHALED"]<-"GLUCOCORTICOIDS"

subset.meds.df$pharm.class.plotting[subset.meds.df$pharm.class.plotting == "TOCILIZUMAB"]<-"IMMUNOSUPPRESSIVES"
                                    
subset.meds.df$study.id<-subset.meds.df$study.id                                    

slow.ct.pts<-rownames(pt.factors.df)[pt.factors.df$Last_Positive_Ct<35 & pt.factors.df$first_last_pos_day>=40]
slow.ct.pts<-slow.ct.pts[!is.na(slow.ct.pts)]


#tiff("fig_jpegs_and_table/2022-03-02_figure_3", units="in",width = 6.87, height #= 9.06, device='tiff', dpi=300)

ggplot(subset(prolonged.pos.df,study.id %in% slow.ct.pts), aes(y=reorder(study.id, first.pos.date), x=Collected, group=study.id)) + 
  geom_jitter( height = 0.01,width = 0.0,size = 2, aes(fill=pos.test), colour="black", pch=21)+
  geom_text_repel(aes(label =Alinity ,angle=75),size=2,max.overlaps = 100,force=1, point.padding=unit(1,'lines'),
                direction = 'y',
                nudge_y = 0.3,
                segment.size=0.2)+
  geom_text_repel(aes(label =Cepheid_RdRp ,angle=75),size=2,max.overlaps = 100,force=1, point.padding=unit(1,'lines'),
                direction = 'y',
                nudge_y = 0.2,
                segment.size=0.2)+
  geom_text_repel(aes(label =M2000 ,angle=75),size=2,max.overlaps = 100,force=1, point.padding=unit(1,'lines'),
                direction = 'y',
                nudge_y = 0.2,
                segment.size=0.2)+
 # geom_text_repel(aes(label =Cepheid_N2 ,angle=75),size=2,max.overlaps = 100,force=1, point.padding=unit(1,'lines'),
               # direction = 'y',
               # nudge_y = 0.2,
               # segment.size=0.2)+
  geom_text_repel(aes(label =Cepheid_E ,angle=75),size=2,max.overlaps = 100,force=1, point.padding=unit(1,'lines'),
                direction = 'y',
                nudge_y = 0.2,
                segment.size=0.2)+
  geom_point(data=subset(subset.meds.df,study.id %in% slow.ct.pts), aes(y=reorder(study.id, first.pos.date), x=START_DATE, group=study.id, colour=factor(pharm.class.plotting)), pch=22,height = 0.1,width = 0.0,size = 1.5)+
  xlab("Time (week) ") +
  ylab("Patients") + 
  theme_bw()+
   theme( panel.grid.major.y = element_line( size=.3, color="darkgrey" ),axis.text.x = element_blank()) +
  #ggtitle("Figure 3:Positive and negative tests prolonged positive patients")+
  labs(fill='RT-PCR Result')+
scale_x_date(breaks = seq(min(repeat.test.df$Collected), max(repeat.test.df$Collected), by = 7), labels=seq(min(repeat.test.df$Collected), max(repeat.test.df$Collected), by =7))+
  theme(axis.text.x = element_text(angle = 90))+
  scale_fill_manual(values=result.colors)+
 scale_shape_manual(values = c(21,22,23,24))+
  xlab("Time (week) ") +
  ylab("Patients")+
  scale_y_discrete(labels=c('796106'="Patient 1","778296"="Patient 2","854418"= "Patient 3","958344"="Patient 4", "87660"="Patient 5", "211382"="Patient 6"))
  
#dev.off()


ggsave(path = "fig_jpegs_and_table/",filename = "2022-03-02_figure_3",
      width =6.87, height =9.06, device='tiff', dpi=300)
```


### Normalization data
-Ct values with test material per reviewer comments
10/20/21  

```{r}

norm.ct.df<-read.csv('2021-10-20_Ct_normalization_formatted.csv')

norm.ct.df$Ct_value[norm.ct.df$Platform=='m2000']<-norm.ct.df$Ct_value[norm.ct.df$Platform=='m2000']+10

norm.ct.df<-norm.ct.df[norm.ct.df$Original_CT != '1000cp/ml',]


ggplot(data=norm.ct.df, aes(y=Ct_value, x = factor(Platform))) +
  #geom_violin(width=.8, adjust=2)+
  geom_boxplot(width=.5, color="black",outlier.size=-1)+
  geom_jitter(height = 0.0,width = 0.24,size = 1)+
ggtitle('Ct values by pool tested on three different days')+
  facet_grid(day~Original_CT)+
  theme(axis.text.x = element_text(angle = 90))+
  stat_summary(
    fun.data = function(x)
      data.frame(y= Inf, label = paste(round(mean(x), 1),'(', round(sd(x),1), ' )')),
    geom = "text",
    aes(group = Original_CT),
    hjust = 0.5, vjust = 1,
    position = position_dodge(width = 1.2),
    size=2.5
  )+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90))+
  xlab("Platform")+
  ylab("Ct value")





ggplot(data=norm.ct.df, aes(y=Ct_value, x = factor(Platform))) +
  geom_violin(width=.7, adjust=2)+
  geom_boxplot(width=.4, color="black")+
  geom_jitter(height = 0.0,width = 0.24,size = 2,aes(shape = day, color=factor(Original_CT)))+ 
  scale_y_continuous(breaks = seq(14,36,2),labels = seq(14,36,2))+
  theme(axis.text.x = element_text(angle = 90))+
ggtitle('Ct values by sample tested across multiple platforms over 3 days')


pdf('2021-10-20_normalization_results_Ct_values.pdf')

ggplot(data=norm.ct.df, aes(y=Ct_value, x = factor(Original_CT))) +
  geom_violin(width=.7, adjust=2)+
  geom_boxplot(width=.3, color="grey")+
  geom_jitter(height = 0.0,width = 0.24,size = 2,aes(shape = day, color=factor(Platform)))+ 
  scale_y_continuous(breaks = seq(14,36,2),labels = seq(14,36,2))+
  theme(axis.text.x = element_text(angle = 90))+
ggtitle('Ct values across platforms by categorical starting material ')+
  theme_bw()


norm.ct.df$all<-'All_samples_tested'

ggplot(data=norm.ct.df, aes(y=Ct_value,x=all)) +
  geom_violin(width=.7, adjust=2)+
  geom_boxplot(width=.3, color="grey")+
  geom_jitter(height = 0.0,width = 0.24,size = 2,aes(shape = day, color=factor(Platform)))+ 
  scale_y_continuous(breaks = seq(14,36,2),labels = seq(14,36,2))+
  theme(axis.text.x = element_text(angle = 90))+
ggtitle('Ct values across platforms')+
  theme_bw()

# test linear model
# ct explained by platform
print(summary(lm(log1p(norm.ct.df$Ct_value)~norm.ct.df$Platform + norm.ct.df$Platform*norm.ct.df$Original_CT + norm.ct.df$Original_CT)))


dev.off()


# simple linear regression ct explained by platform:
# not a good predictor

summary(lm(formula = log1p(norm.ct.df$Ct_value) ~ norm.ct.df$Platform))
```

### Summary table for normalization
```{r}

norm.summary.df<-as.data.frame(expand_grid(unique(norm.ct.df$Platform), unique(norm.ct.df$Original_CT)))
colnames(norm.summary.df)<-c('Platform','Pool_range')

norm.summary.df$mean<-NA
norm.summary.df$sd<-NA

for(platform in norm.summary.df$Platform){
  for(ct in norm.summary.df$Pool_range){
    
    current.mean<-mean(norm.ct.df$Ct_value[norm.ct.df$Platform ==platform & norm.ct.df$Original_CT==ct])
    current.sd<-sd(norm.ct.df$Ct_value[norm.ct.df$Platform ==platform & norm.ct.df$Original_CT==ct])
    
    norm.summary.df$mean[norm.summary.df$Platform==platform & norm.summary.df$Pool_range==ct]<-current.mean
     norm.summary.df$sd[norm.summary.df$Platform==platform & norm.summary.df$Pool_range==ct]<-current.sd
    
    
  }
}

write.xlsx(norm.summary.df[order(norm.summary.df$Pool_range),], file='fig_jpegs_and_table/table_S1.xlsx')

```

